---
title: "sensi_an_rotations1 : start of sensitivity analysis for rotations"
output: pdf_document:
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---


```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
  library(rotations)
  library(ggplot2)

  outFolder <- "C:\\Dropbox\\resistanceResults\\rotations\\"
```

Starting question : is there any benefit to rotations when there are no costs and no migration to an untreated area ?

Test 1. No refugia (i.e. Coverage=100%) or fitness costs. Would expect all policies to be equivalent.

Test 2. Dominance of resistance and fitness costs both set to 0.5.
I suspect one of the advantages of rotations (and the presence of refugia) is that they maintain lower resistance allele frequencies which slows the spread of recessive resistance alleles. I suspect/guess/intuit that if we make them semi-dominate these effects should disappear and all policies should be the same. I’m probably wrong but it’s a nice test…..

```{r, eval=TRUE}

#start small & build it up
#maybe create an equiv. of resistance::sensiAnPaperPart.r later

# example for 1 run
# linputs <- list(max_gen = 50, n_insecticides = 2)
# do.call(run_rot, linputs)
# dfinputs <- as.data.frame(linputs)

# So one way of doing is to create an linputs list for each run & pass to run_rot()
# so run_rot() only ever runs for 1 scenario
# in contrast resistance::runModel2() accepted an input matrix & would run for multiple scenarios
# not sure which is better

n_scenarios <- 5

df_res_all <- data.frame()

for(i in 1:n_scenarios)
{
  # expo_hi 0.1 – 0.9
  # eff 0.3 - 1
  linputs <- list()
  
  linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
  linputs$eff <-      runif(1, min=0.3, max=1)
  
  # run one scenario
  dfres <- do.call(run_rot, linputs)
  
  # summarise results per scenario
  # NOTE this is also currently calc at end of run_rot() prob better to calc there
  # and put in the output dfres, but its then repeated per 
  # TODO first AG fix this maybe go for non dplyr solution
  # if I do calc in run_rot then just find value at first row for each gene
  # Error in dplyr::filter(active_or_refuge == "active") : 
  # object 'active_or_refuge' not found
  res <- dfres %>%
  # only assess in control areas not in refugia  
  dplyr::filter(active_or_refuge=='active') %>%
  group_by(resist_gene) %>%
  # for all insecticides in all generations  
  # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
  # summarise(mean_gens_under50 = mean(gens_under50)) %>%
  # just for deployed insecticides 
  summarise(gens_dep_under50 = sum(resistance < 0.5 &
                                   #finds insecticide in use = this one
                                   #TODO in_use is also calculated in rot_plots, but set to 1.05 for the plots
                                   resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
  summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
  unlist()
  
  # add results onto inputs list
  linputs <- c(linputs, res)
  
  # change list to dataframe
  df_in_out <- as.data.frame(linputs)
  
  # rbind this scenario inputs & results onto previous
  # not a good speed or memory way of doing ... but a start
  df_res_all <- rbind(df_res_all, df_in_out)
  
}


```

<!-- eval=FALSE to stop slow code being executed after 1st time --> 
```{r, eval=FALSE}

  # this is how it was done for resistance paper1
  # nScenarios <- 10  
  # ## do model runs
  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
  # ## save results objects as rda for re-use
  # save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))

     
  
```  
